\def\cdca{\textsc{DOC}}


\makeatletter
\def\ednoteshape{\sf}
\newif\ifshowednotes\showednotesfalse
\def\edshownotes{\showednotestrue}
\DeclareOption{show}{\showednotestrue}
\DeclareOption{hide}{\showednotesfalse}
\ProcessOptions

\newcounter{ednote}
\def\ed@foot#1#2#3% text, type, label
{\def\test{#3}\def\empty{}\footnotetext[\value{ednote}]%
{{\sc{#2}\if\test\empty\else\label{ed:#3}[{#3}]\fi:} \ednoteshape #1}}
\def\ed@note#1#2#3% text, type, label
{\addtocounter{ednote}{1}\message{#2!}%
\ifshowednotes%
\footnotemark[\arabic{ednote}]\ed@foot{#1}{#2}{#3}%
\marginpar{#2(\arabic{ednote})}%
\fi}
\newcommand{\ednote}[2][]{\ed@note{#2}{EdNote}{#1}}
\newcommand{\issue}[2][]{\ed@note{#2}{Issue}{#1}}
\newenvironment{newpart}[1]% text
{\addtocounter{ednote}{1}\edef\new@number{\theednote}\message{New Part!\new@number}
\ifshowednotes\ed@foot{#1}{New Part}{}\marginpar{BegNP(\new@number)}\fi}
{\ifshowednotes\marginpar{EndNP(\new@number)}\fi}

\def\ednotemessage{\ifnum\value{ednote}>0\typeout{}%
\typeout{There are still \arabic{ednote} EdNotes and Issues to resolve!}%
\typeout{}\fi}
\makeatother
%% a bit wider
%\advance\evensidemargin-.8in
%\advance\oddsidemargin-.8in
%\advance\textwidth1.6in
% Vertical space mods from Mike's old 2.09 version.
% Separate paragraphs, and subvert \item to leave less space!
%
\parskip=0.5\baselineskip
\parindent=0pt

\documentclass{euproposal}

\usepackage{paralist,enumerate,url}

\begin{document}
\setcounter{part}{2}% part B




\title{Realizing the Potential of Semantic  Markup for Mathematics\\
  \normalfont\large Future and Emerging Technologies Open Coordination
  Action}

\date{Date of Preparation: \today}
\author{Coordinator: Professor Michael Kohlhase\\
  School of Engineering \& Sciences\\
  International University Bremen, \\
  Campus Ring 12, D-28759 Bremen, Germany\\ 
\texttt{m.kohlhase@iu-bremen.de}\\
tel: +49 421 200 3140 \\ fax: +49 421 200 3103} 


\titlepage 

\maketitle
\showednotestrue
\newpage
\tableofcontents

\setcounter{chapter}{0}

\section{Proposal Summary}

\textbf{Realizing the Potential of Semantic Markup for Mathematics}\\
\textsc{Acronym:} \\


\subsection{Strategic objectives}

The two key objectives of this proposal are:
\begin{itemize}
\item To foster and extend the use of OpenMath and MathML within the
  European scientific community.
\item To embed these technologies into the semantic web by providing
  views of them grounded in suitable ontologies.  This will enable
  them to be used by applications based on RDF and OWL.
\end{itemize}


\subsection{Abstract}

There is a critical mass of activity developing around the area of
semantic markup for mathematics embracing the fields of publishing,
e-science, and mathematical software.  Much of this activity is or has
been sponsored by the EU (OpenMath, OpenMathTN, MONET, WebALT,
LeActivemath, MOWGLI, Calculemus, ...), some is sponsored by national
agencies (MathBroker, GENSS, ...), some by multi-national consortia
(W3C's MathML Interest Group, The OpenMath Society, ...), and some by
private enterprise (NAG Ltd., Springer-Verlag, Design Science Inc,
Stilo Technology Ltd. ...).  Coordination of these activities and the
maintenance of agreed standards will have tremendous long-term
benefits for the scientific and technical community in Europe and the
wider world.




\chapter{Objectives}\label{cha:object}

In the 21st Century one of the great challenges facing the scientific
community will be to manage increasingly large bodies of
digitally-encoded information.  To allow this information to be
\emph{re-used} in different contexts, and to be \emph{processed} by
software systems,
%amc: replaced assistants by systems to avoid too much of our own idiom
it will need to be encoded using appropriate semantic markup.  As a
result there has in recent years been an explosion of interest in the
semantic web and related technologies.

In the OpenMath Project~\cite{OM-EU} and the Thematic
Network~\cite{OM-TN} that followed it, the European Union supported
the development of OpenMath and MathML~\cite{}, two complementary standards
for content and presentation markup of mathematics.  The original
intention was that these would be used to exchange mathematical
objects between software systems, facilitate the development of
``active'' texts where the mathematical components could be extracted,
manipulated and checked.  As such, these technologies have been a
great success and are now beginning to be used in a wide range of
areas.  However now we see the potential to combine our existing
markup mechanisms with emerging semantic web technologies to allow for
digital mathematical content to be managed, searched, indexed,
validated etc.~in a mathematically meaningful way.

The two key objectives of this proposal are:
\begin{itemize}
\item To foster and extend the use of OpenMath and MathML within the
  European scientific community.
\item To embed these technologies into the semantic web by providing
  views of them grounded in suitable ontologies.  This will enable
  them to be used by applications based on RDF and OWL.
\end{itemize}


\chapter{Relevance to the objectives of FET Open}
\label{cha:relevance}

Several research groups in Europe are currently using and developing
semantic markup for mathematics. Their activities range from
publishing, to e-science on the grid, and mathematical software
development. In the past, EU-sponsored projects in this area,
(OpenMath, OpenMathTN, MONET, MOWGLI, Calculemus) have developed new
using emerging web technologies and standards for the electronic
communication of mathematics.  Current EU projects (WebALT,
LeActivemath), national agencies (MathBroker, GENSS, ...),
multi-national consortia (W3C's MathML Interest Group, The OpenMath
Society, ...), and some private enterprise (NAG Ltd., Springer-Verlag,
Design Science Inc, Stilo Technology Ltd., Math4More ...) build upon
and exploit such technologies. Integration of all these related
activities is crucial for the advancement in the emerging area of
semantic web technologies. This coordination action will monitor and
align the on-going international efforts that are extending OpenMath
in emerging markup languages for their activity to promote a common
standard.





\chapter{Potential Impact}\label{cha:impact}



Coordination of the wide range of activities that involve semantic
markup of mathematics, the maintenance of agreed standards, the
establishments of expert groups and usage guidelines will have
tremendous long-term benefits for the scientific and technical
community in Europe and the wider world.




\chapter{Consortium and Project Resources}\label{cha:consortium}

\begin{description}
  
\item[NAG Ltd] NAG is Europe's largest vendor of mathematical
  software.  It is involved heavily in the W3C Math Working Group,
  providing one of the current co-chairs.  NAG uses semantic markup to
  tailor documentation for specific user communities, and to generate
  software components from abstract specifications.  In addition to
  coordinating the Thematic Network is ran the website and developed a
  range of validation tools for OpenMath documents and Content
  Dictionaries.  It also led the development of the OpenMath~2
  Standard.
  
\item[University of Bath] The Computing Department at the University
  of Bath has a long history of involvement in Computer Algebra, and
  until recently was part of the School of Mathematical Sciences.  It
  coordinated the development of Content Dictionaries during the
  former projects: OpenMath and OpenMath Thematic Network.
  
\item[Stilo Technology Ltd] Stilo is an SME providing XML-based
  services in a broad range of areas.  It is involved in a number of
  W3C activities (including MathML) and was heavily involved in the
  development of OpenMath~2.
  
\item[INRIA Sophia-Antipolis] INRIA is one of the three original W3C
  host institutions and the group at Sophia-Antipolis has been
  involved in the Math Working group from its inception.  They also
  have a long history in the development of mathematical software and
  have been a key developer within the MONET project.
  
\item[University of St Andrews] St Andrews coordinates development of
  the GAP algebraic software system, which can read and write
  OpenMath.
  
\item[Technical University of Eindhoven] The group at Eindhoven has
  been involved in the development of electronic mathematical
  textbooks, notably \emph{Algebra Interactive} for some years, and
  are using OpenMath to communicate between user interfaces and a
  range of mathematical engines.  They contributed a significant
  number of Content Dictionaries during the OpenMath Thematic Network
  project and have played a key design and development role for the
  MONET project. 
  
%  \item[Springer Verlag] Springer is interested in electronic and
%  interactive publishing, and have collaborated with Eindhoven in the
%  development and publication of \emph{Algebra Interactive}.
 
\item[University of Nice] The M@INLINE Project at the University of
  Nice is interested in e-learning and the development of tools for
  delivering interactive courses in Mathematics.  They have developed
  various OpenMath tools including an editor and a rendering agent.
  
\item[Konrad-Zuse Zentrum fur Informationstechnik] ZIB is one of the
  main developers of the Reduce Computer Algebra System.

\item[Research Institute for Symbolic Computation] RISC is one of the
  largest research centers for Computer Algebra in the world.  It
  played a substantial part in the development of the OpenMath~2
  Standard and of the MONET protocols.
  
\item[German Research Centre for Artificial Intelligence] Based at the
  DFKI, the ActiveMath project is developing technology for
  e-learning.  They make substantial use of OMDoc, a mechanism for
  representing mathematical documents which is built on OpenMath.
  
\item[University of Helsinki] The University of Helsinki is also
  interested in e-learning and, through its collaboration with Florida
  State University, has developed a substantial body of material for
  teaching undergraduate calculus.
  
\item[University of Cologne] Cologne is interested in the theory
  behind semantic markup for mathematics, and provided in-depth and
  thorough critiques of drafts of both MathML~2 (second edition) and
  OpenMath~2.
  
\item[International University of Bremen] Bremen is the lead developer
  of the OMDoc system for representing complete mathematical documents
  in a machine-processable form.  They are involved in the W3C Math
  Working group and were involved in the development of OpenMath~2.

\end{description}


\chapter{Project Management}\label{cha:management}
\begin{newpart}{adapted from the old OpenMath project via the MKM proposal; this
    needs to be discussed}   
  The proposers intend to adapt the management systems that have
  worked successfully in the OpenMath project and other networks to
  which several of them belonged. The following key individuals will
  be involved in the management of the network.
\begin{enumerate}
\item Project Chairman. \ednote{Michael?}
\item Project Manager. \ednote{Olga?}
\item Site Leaders. Each site will nominate one site leader.
\item Work Package Leaders. The Project Manager will nominate one for
  each Work Package (with himself as the leader of Work Package 0)
  from the site that has taken lead responsibility for drafting this
  Work Package. These may be changed by the Project Board.
\item Task Leaders. These are nominated by the Project Manager on the
  advice of the Work Package Leaders.
\end{enumerate}
The first three categories constitute the Supervisory Board, which
will meet at each of the annual plenary meetings.  It will have the
ultimate responsibility for managing the project. Work Package and
Task Leaders, if not already members, may be invited for the
discussion of particular issues, but will not be allowed to vote.
   
Day-to-day management, and budgetary allocation, will be in the hands
of a smaller Management Board: Project Chairman, Project Manager and
representatives of various interest groups.

In practice, and based on experience in other EU projects to which
many of the members have belonged, it is generally possible to manage
such a coordination action very largely by e-mail, though it is
probable that conference calls between the Management Board will be
necessary as well.
  
The Management Board, as advised by the Project Manager, will have the
responsibility for the management of the project within the contract
between the Commission and the Coordinator.
  
If a conflict arises, or is noticed by the Project Manager, he may
attempt to resolve it himself. If he does not, or if this fails, then
he refers the conflict to the Project Chairman. The latter may also
attempt to resolve it himself. If he does not, or if this fails, then
the Project Chairman refers it to the Management Board. He can either
call a special meeting of the Board (with at least two weeks notice
given to all members by fax and e-mail) or wait until the next
scheduled meeting of the Board, depending on the timing and the
importance of the conflict.  At the meeting, each member to the
conflict may state their case. The Board will then vote by secret
ballot. In the event of a tie, the Chairman will have a casting vote.

It is envisaged that, under the Management Board, there will be
Committees responsible for each of
\begin{itemize}
\item Integrating Activities
\item The Joint Programme of Activities --- all Work Package leaders
  would belong to this Committee;
\item Training and Dissemination Activities.
\end{itemize}
However, the full structure of these Committees would have to await
the first meeting of the Supervisory Board to approve the Management
Board's recommendations.
\end{newpart}

\chapter{Workplan}\label{cha:workplan}



\section{Introduction}

The ability to use computers for performing mathematical tasks is
undoubtedly one of the key factors behind recent progress in science,
engineering, and economics.  New technology such as the computational
grid and the semantic web could provide the infrastructure which will
allow researchers to make even greater leaps forward by enhancing
collaboration and providing instant access to massive computational
resources.  On a smaller scale, web service technology holds out the
promise of better interaction between mathematical applications within
a particular community.

These visions rely on the ability to move data and mathematical
objects between pieces of software, and this in turn requires a
mechanism for describing the meaning of those objects.  This issue was
addressed in the Fifth Framework OpenMath and OpenMath Thematic
Network projects, in cooperation with the Worldwide Web Consortium's
(W3C) Math Working Group.  The result is two complementary
standards~--- OpenMath and MathML~--- which allow mathematical objects
to be transmitted and processed by computers in a mathematically
meaningful way, as well as rendered in the current generation of web
browsers.

The coordination action focuses on strengthening OpenMath support for
mathematical communication in the scientific, engineering, and
economics communities, and on supporting European involvement in the
newly-chartered W3C Math Interest Group.  The coordination action will
sponsor activities such as

\begin{enumerate}[(i)]
\item the enhancement and strengthening of the community of OpenMath
  users;
\item further evolution of the OpenMath vocabulary;
\item tailoring OpenMath for use as a technology within the semantic
  web;
\item the development of web and grid services which communicate via
  OpenMath;
\item the development of tools for OpenMath developers.
\end{enumerate}


\subsection{Strengthening the OpenMath Community}\label{community}


As a result of the previous projects there is now a broad community of
people using OpenMath in a variety of ways.  Activities in the areas
of publishing, e-learning, web services and conventional mathematical
software are all in progress.  The main aim of this task is to provide
a basic infrastructure in which these groups can collaborate. This
includes providing web-based and hands-on training, supporting mailing
lists and organizing regular workshops.

This task is at the core of the coordination action since it aims at
strengthening the user base of OpenMath by promoting Content
Dictionaries (CDs), by dissemination and training, setting up an editorial
board for CDs, expert consulting, easy to follow and good practice
guidelines. The OpenMath web site at \texttt{http://www.openmath.org}
will collect the results produced during this action's lifetime, and
maintain and enhance the available content as a central resource for
the OpenMath community. 


% I think that it would be good to have a section after the introduction
% describing the existing OpenMath Community.  This could be structured as
% follows:
% - Mathematical Software (NAG, ORCCA/Maple, Axiom, Cocoa ...)
% - Mathematical Publishing (OMDoc, Algebra Interactive, NAG)
% - E-Learning (Eindhoven, LeActiveMath, Helsinki)
% - Semantic Web (MONET, Stilo)
% - Web Services (MONET, ORCCA/Maple)
% Obviously the more groups and the wider the coverage of topics the
% better.  We might want to think about who we could bring "into the fold"
% as well.



\subsubsection{Mathematical Software}\label{sec:msw}
% - Mathematical Software (NAG, ORCCA/Maple, Axiom, Cocoa ...)

The original and primary target user group for OpenMath was the
community of mathematical users and developers. OpenMath has been
developed with the intent of supporting easy interfacing of
mathematical software packages with different yet compatible
functionalities.

Examples of computer algebra systems that speak OpenMath include
libraries such as Aldor, general purpose systems such as Maple,
REDUCE, and AXIOM, special purpose systems such as GAP and CoCoA.
Furthermore, OpenMath has been taken up as an communication standard
by (semi)-automated theorem provers and program verification systems
through the OMDoc format in the OMEGA, TPS, OTTER, and PVS.

%what to do to strengthen this community 
The coordination action intends to strengthen this community by adding
more visibility to the OpenMath compliant software via the web page,
by collecting and disseminating guidelines for implementors, by active
support for the upgrade to OpenMath version 2.0, by the development
and exchange of Content Dictionaries and Phrasebooks.

\subsubsection{Mathematical Publishing}\label{sec:mpubl}
% - Mathematical Publishing (OMDoc, Algebra Interactive, NAG)

The mathematical publishing community broadly includes publishing
companies of scientific literature, software companies distributing
documentation and training material, down to the researchers that
distribute results and lectures on-line.

Publishers of mathematical books are interested in the inclusion of
OpenMath in electronic versions of texts because it enhances
interactivity: with a mouse click, a mathematical object can be
transferred to a computational program or to visualization software.
The same reason holds true for OpenMath to be incorporated in other
publishing activities: the mathematics in published material becomes
truly interactive whether it is demonstrating usage of third-party
algorithms or giving hands-on examples to students.

The publishing community can follow the examples set by Algebra
Interactive (TU Eindhoven/Springer), by OMDoc activities (IUB, DFKI,
Carnegie Mellon University) and by NAG (hyper-textual documentation).

%what to do to strengthen this community
The coordination action intends to strengthen this community by
cooperating with activities such as the Digital Library of
Mathematical Functions (DLMF) developed at the National Institute of
Standards and Technology (NIST)
\footnote{\url{http://www.openmath.org/meetings/helsinki2004/Miller/slides.pdf}},
collect tools that simplify inclusion of OpenMath in published
material, include OpenMath in accepted formats such as DocBook.



\subsubsection{E-Learning}\label{sec:e-learn}
% - E-Learning (Eindhoven, LeActiveMath, Helsinki)
Novel usage of OpenMath comes from the e-learning community. To
e-learning developers, OpenMath offers the advantage of being a rather
simple language for the representation of mathematics that
standardizes the various notations used in computational software
systems. Authors can produce electronic lectures in a format that can
be visualized on the screen, read out automatically or printed. The
same source may be used to produce handouts of specific lesson-trails
or dynamically adapted to the user profile. The mathematical
expressions can be directly fed to a variety of software packages and
to mathematical web services.

LeActiveMath (IST-507826) is a EU project targeting the development of
a third generation e-learning system that is Language-Enhanced, User
Adaptive, and provides Interactive eLearning for Mathematics. OpenMath
is the internal XML format chosen for the representation of the
mathematical content.  The Helsinki Learning System provides tools to
create multilingual adaptive exercise databases that support learning
in any discipline, currently it covers a typical single variable
calculus course in mathematics. The prospective EU project WebAlt
intends to extend the system with OpenMath support.

\begin{newpart}{MiKo: as requested}
  Outside the EU, the ``Open Learning Initiative'' and the ``Course Capsules
  Project'' at Carnegie Mellon University have built on the OpenMath/OMDoc format,
  extending it to courses in Computer Science and Philosophy, adding authoring
  support in the {\LaTeX}, MS-PowerPoint, and Mathematica workflows.
\end{newpart}

The coordination action intends to strengthen this community by
increasing dissemination in relevant conferences (IMACS-ACA,
technology conference in mathematics), by producing easy-to-follow
self-contained examples beyond the mathematical field, by enlarging
the number of authors to scientific and engineering fields.

\subsubsection{Markup Languages}\label{sec:ml}

The user community of markup languages is growing every day and has
the potential of becoming a large user group for OpenMath, should it
be adopted as the markup language of mathematics inside scientific,
engineering and commercial markups (see the recent workshop
\url{http://scimarkuplang.comm.nsdlib.org/cgi-bin/wiki.pl}).

MathML is a format for describing both the presentation and content of
mathematical objects, and is designed to support OpenMath as a content
descriptor.  The W3C Math Working group was recently dissolved, after
completing an editorial revision of MathML.  A new Math Interest Group
has been set up by W3C to support users of MathML, and an important
aspect of this task will be to support European involvement in it.
One of the members of this Consortium, NAG Ltd, provides one of the
co-chairs of the Math Interest Group.

CML (Chemical Markup Language) and STMML (Scientific, Technical and
Medical Markup Language), RoboML (Robotic Markup Language), are at
present a few examples of markup languages that only support OpenMath
indirectly through MathML.  Public awareness will increase by crafting
examples in these languages which are representable in MathML only by
resorting to OpenMath Content Dictionaries.

\subsection{Evolving the OpenMath Vocabulary}\label{sec:OCD}

The mathematical knowledge is provided in OpenMath by means of
symbols, defined in ``OpenMath Content Dictionaries'' (CDs).  OpenMath
Content Dictionaries make available mathematical definitions and
symbols in an extensible, machine-readable way.

For OpenMath to become widespread it is important to have a critical
mass of CDs. This task is centered on identifying and producing CDs
for those areas of mathematics which are not yet supported by either
MathML or OpenMath. The mathematical content will be represented in
CDs and these CDs will be implemented and tested in one or more
software packages as to demonstrate their usefulness.

Special attention will be devoted to CDs linked to 
\begin{itemize}
\item computational software packages which decides to become part of
  the OpenMath community by for instance implementing an OpenMath
  \texttt{export} facility (Maple, Axiom, Singular, CoCoA, Macaulay2,
  Cinderella, Wiris, FORM, \ldots)
\item projects that use and/or extend OpenMath (LeActiveMath, MONET,
  Algebra Interactive, \ldots)
\item computationally relevant fields ranging from the Digital Library
  of Mathematical Functions at NIST to specific case studies such as
  the problem of matrix-group recognition (solvable by GAP and Magma)
  or dynamic geometry (handled by Cinderella and Geometric Discovery
  (see {\url{http://193.146.36.49/discovery}}). Special effort will be
  devoted to producing CDs for fields new to OpenMath such as
  financial mathematics, physical and engineering applications.
\end{itemize}
\begin{newpart}{MiKo: to say that the preconditions have changed favorably}
  With the OpenMath2 standard accomplished in the OpenMathTN, we have
  generalized the content dictionary format, so that content
  dictionaries e.g. in OMDoc form can now be self-documenting entities
  that also develop the necessary mathematical material. In this way,
  e.g. the DLMF can serve as a web-accessible tome of information
  about special functions as well as their content dictionary, which
  fixes the semantics of OpenMath objects referencing them. This
  removes the need for separate content dictionaries and alleviates a
  major knowledge management hassle and roadblock to content
  dictionary adaption.
\end{newpart}

\subsection{OpenMath inside the Semantic Web}

W3C is developing tools that will support the next generation
\emph{semantic} web where information is encoded in such a way that
its meaning is unambiguous and where software agents can process and
reason with it on behalf of human users.  In the semantic web vision,
a piece of mathematics is identified with its meaning, the content,
which must be uniquely understood and unambiguous.  OpenMath has
always focused on the problem of creating a representation of
mathematics which is notation independent and carries the mathematical
content, even before the World Wide Web existed in the way we
experience it now. With the recent development of technologies, like
OWL and RDF, meant to underpin a semantic web, OpenMath can play an
important role as the ontological framework of mathematics. The MONET
project (IST-2001-34145) produced results which have already started
to build ontologies that bridge OpenMath to RDF and OWL.  In
particular, MONET experimented with using OpenMath inside the Web
Ontology Language so that a software component could make decisions by
reasoning about OpenMath objects.

This task will continue to investigate the integration of OpenMath
into the semantic web.  One aspect of this is to allow OpenMath
objects to be manipulated by semantic web tools, as was done in MONET.
Another is to develop and further refine the concept of Content
Dictionaries so that the information contained within them can be
processed by sophisticated tools such as proof assistants and theorem
provers.

\subsection{Web and Grid Services}


% Rewritten from MKM proposal

Web Service technology is a fundamental development towards
distributed collaboration on the computational grid (often called
\emph{e-science}).  The Web Service paradigm can be used in the
mathematical context to offer controlled access to literature
information, e.g. journal articles, or to databases, e.g.  integral
tables, or to offer access to computational resources, e.g.
differential equation solvers, theorem provers, statistical packages.
Evidence from the EU MONET project (IST-2001-34145) suggests that true
advertisement and discovery of mathematical services requires building
and maintaining a significant amount of mathematical infrastructure
beyond that offered by the web service technologies used in industry.
The MONET framework uses a layered approach for the mathematical parts
of the service description, the \emph{Mathematical Service Description
  Language} (MSDL) in which the mathematical fragments of the
descriptions heavily rely on OpenMath.  The ultimate goal is for a
user to be able to state requirements specifically (``solve
$\int_0^\pi{}\sin(x)\,dx$'') or generically (``find me an available
service which performs definite integration and costs nothing for me
to use'') and a mathematical broker to find the suitable services that
match the request.

GENSS (Grid-Enabled Numerical and Symbolic Services), a joint project
between the Universities of Bath and Cardiff, addresses the
combination of Grid computing and mathematical Web services, and their
extension to deliver mathematical problem analysis, and the code and
the resources to compute the answers, using a common open agent-based
framework.. The project builds on the work on Mathematical Web
Services done in the MONET project.

%what to do to strengthen this community
The coordination action intends to strengthen the usage of OpenMath in
e-science by building a MONET expert group that develops and
maintains standards for mathematical service descriptions. To foster
collaboration, this task will setup a public registry of mathematical
services along with the necessary Problem and Algorithm Description
Libraries.  The group will monitor the new web service technologies
developed by Worldwide Web Consortium and OASIS to ascertain their
compatibility with the universe of mathematical applications.


\subsection{Tools Development}\label{tools}

This task realizes the need for OpenMath tools with special emphasis
on tools that facilitate the creation and usage of CDs. Activities
that develop editors, stylesheets, validation mechanisms, archiving
and searching facilities fall under this task.

Tools such as the following will receive special attention by the
coordination action: 
\begin{itemize}
\item the OpenMath/OMDoc editor under development within the European project
  LeActiveMath
\item the MBase knowledge base for archiving and searching Content
  Dictionaries symbols, mathematical properties, signatures.
\begin{newpart}{MiKo: some more, need even more}
\item The editor CPoint (OpenMath/OMDoc in MS PowerPoint) under development at
  University Bremen
\item The OpenMath/OMDoc extension for the GNU {\TeX}macs editor under development
  at Saarland University.
\item The S{\TeX}/{\LaTeX}ML workflow for semantic annotation and conversion of
  {\LaTeX} to OpenMath under development at NIST and IUB.
\item second-generation presentation systems for OpenMath: in particular modular
  formats for including notation information into content dictionaries and their
  integration into OpenMath presentation- and translation systems.
\end{newpart}
\end{itemize}
\section{Work Plan}

The work plan of ???? is centered around eight work packages, one for
each area identified in the previous section plus one for management.
As the work packages are relatively orthogonal, they can basically run
in parallel, and will run for the whole duration of the
project.\ednote{adapt and extend, where necessary} The following two
tables give an overview over the work packages and their
deliverables.\ednote{please volunteer (evenly) for lead contrator
  roles. }

\begin{tabular}{|>{\centering}p{20mm}|l|
       >{\centering}p{20mm}|>{\centering}p{20mm}|>{\centering}p{20mm}|}
\hline
\bfseries Work\-package No&\bfseries Work package Title&
\bfseries Lead Contractor&
\bfseries Start month&
\bfseries End month\tabularnewline\hline
WP 0& Management                         & IUB  & 1 & 36 \tabularnewline\hline
WP 1& OpenMath in Software               & ???   & 1 & 36 \tabularnewline\hline
WP 2& OpenMath in Publishing             & ???   & 1 & 36 \tabularnewline\hline
WP 3& OpenMath in E-Learning             & ???   & 1 & 36 \tabularnewline\hline
%WP 4& OpenMath in E-Science              & IUB   & 1 & 36 \tabularnewline\hline
WP 4& Evolving the OpenMath Vocabulary   & ???   & 1 & 36 \tabularnewline\hline
WP 5& OpenMath and the Semantic Web      & ???   & 1 & 36 \tabularnewline\hline
WP 6& OpenMath in Web- and Grid Services & NAG?  & 1 & 36 \tabularnewline\hline
\end{tabular}

The next table gives an overview over the deliverables of the work
packages. Note that the times of deliverables after month 18 are
estimations and may change as the work packages progress.

\begin{longtable}{|l|l|p{1cm}|l|l|}\hline
\bfseries Number &
\bfseries Deliverable\slash milestone title&
\bfseries due (mo)&
\bfseries Nature&
\bfseries Dissemination level\tabularnewline\hline\hline
D1.1 & ?????????????           & ?? & ?? & public \tabularnewline\hline
\end{longtable}
\vfill\eject

\section{Work Packages}
\newenvironment{workpackage}[5]% duration, number, start, participants, title
{\begin{center}{\Large\bf Workpackage description\par
   WP {#2}: {#5} ({#1} Months)  \vspace{1cm}}
  \fbox{\parbox[t]{\textwidth-15pt}{\raggedright
\makebox[.4\hsize][l]{\textbf{Workpackage number:} WP{#2}}%
\makebox[.6\hsize][l]{\textbf{Start date or start event:} {#3}}\\
\textbf{Participants:} {#4}}}}
{\end{center}}
\newsavebox{\fmbox}
\newenvironment{fmpage}[1]{\begin{lrbox}{\fmbox}\begin{minipage}[{#1}}{\end{minipage}\end{lrbox}\fbox{\usebox{\fmbox}}}
\newenvironment{wpbox}[1]%
{\begin{lrbox}{\fmbox}\begin{minipage}[t][4cm]{\textwidth-15pt}{\bf{#1}\qquad}}%
{\end{minipage}\end{lrbox}\fbox{\usebox{\fmbox}}}

%%%%%%%% WP 0 
\begin{workpackage}{36}{0}{Start of Project}{IUB}{Management}
\begin{wpbox}{Objectives}
  The aim of this work package is to manage the project, liaise with
  the Commission and the reviewers appointed by them, to respond to
  external events, and to ensure that the aims and objectives of the
  overall project are met.
\end{wpbox}
\begin{wpbox}{Description of Work}
Not applicable.
\end{wpbox}
\begin{wpbox}{Deliverables}
  Not applicable (although it will produce internal guidelines and
  regular management reports as well as the final report, as required
  under the contract).
\end{wpbox}
\begin{wpbox}{Milestones and expected result}
\end{wpbox}
\end{workpackage}
\newpage

%%%%%%%%%%%%%% WP 1
\begin{workpackage}{36}{1}{Project Start}{UNIGE,TUE}{OpenMath in Software}
\begin{wpbox}{Objectives}
  To update the software to be compliant with version of OpenMath2.0
  and at the same time build the expertise and guidelines on how to
  implement OpenMath in software.
\end{wpbox}
\begin{wpbox}{Description of work}
  ????????
\end{wpbox}
\begin{wpbox}{Deliverables\\}
  {\bf D1.1} Report on Current Status of OpenMath-compliant Software\\
  {\bf D1.2} Prototype Release of OpenMath2.0-compliant Software\\
  {\bf D1.3} Final OpenMath2.0-compliant Software Upgrade accompanied
  by documentation on the upgrade
\end{wpbox}
\begin{wpbox}{Milestones and expected result}
\begin{itemize}
\item (Month 12) ????
\item (Month 24) ???
\item (Month 36) ???
\end{itemize}
\end{wpbox}
\end{workpackage}
\newpage
%%%%%%%%%%%%%% WP 2
\begin{workpackage}{36}{2}{Project Start}{UNIPI, TUE}{OpenMath in Publishing}
\begin{wpbox}{Objectives}
????????????? 
\end{wpbox}
\begin{wpbox}{Description of work}
  ???????? 
\end{wpbox}
\begin{wpbox}{Deliverables\\}
{\bf D2.1} \\
{\bf D2.2} \\
{\bf D2.3} 
\end{wpbox}
\begin{wpbox}{Milestones and expected result}
\begin{itemize}
\item (Month 12) ????
\item (Month 24) ???
\item (Month 36) ???
\end{itemize}
\end{wpbox}
\end{workpackage}
\newpage

%%%%%%%%%%%%%% WP 3
\begin{workpackage}{36}{3}{Project Start}{UH, DFKI, TUE, KOLN}{OpenMath in E-Learning}
\begin{wpbox}{Objectives}
  Coordinate the usage of OpenMath in e-learning frameworks by
  ....exchanging personnel, visiting respective project meetings,
  comparison study?
\end{wpbox}
\begin{wpbox}{Description of work}
  ????????
\end{wpbox}
\begin{wpbox}{Deliverables\\}
{\bf D3.1} \\
{\bf D3.2} \\
{\bf D3.3} 
\end{wpbox}
\begin{wpbox}{Milestones and expected result}
\begin{itemize}
\item (Month 12) ????
\item (Month 24) ???
\item (Month 36) ???
\end{itemize}
\end{wpbox}
\end{workpackage}
\newpage
%%%%%%%%%%%%%% WP 4
\begin{workpackage}{36}{4}{Project Start}{BATH}{OpenMath in Markup Languages}
\begin{wpbox}{Objectives}
  To encourage the adoption of OpenMath/MathML in markup languages
  that need to represent mathematical objects.  To continue liaison
  between OpenMath and MathML.
\end{wpbox}
\begin{wpbox}{Description of work}
  ????????
\end{wpbox}
\begin{wpbox}{Deliverables\\}
{\bf D4.1} \\
{\bf D4.2} \\
{\bf D4.3} \\
\end{wpbox}
\begin{wpbox}{Milestones and expected result}
\begin{itemize}
\item (Month 12) ????
\item (Month 24) ???
\item (Month 36) ???
\end{itemize}
\end{wpbox}
\end{workpackage}
\newpage

%%%%%%%%%%%%%% WP 5
\begin{workpackage}{36}{5}{Project Start}{TUE, NAG, IUB, BATH}{Evolving the OpenMath Vocabulary} 
\begin{wpbox}{Objectives}
???????
\end{wpbox}
\begin{wpbox}{Description of work}
  ????????
\end{wpbox}
\begin{wpbox}{Deliverables\\}
{\bf D5.1} Report on Status Quo on OpenMath Content Dictionaries\\
{\bf D5.2} Content Dictionaries for Computational Software Packages\\
{\bf D5.3} Final Report on the state of the OpenMath Vocabulary\\
\end{wpbox}
\begin{wpbox}{Milestones and expected result}
\begin{itemize}
\item (Month 3) Assessment of the current state (Deliverable D5.1)
\item (Month 24) First draft for CDs for computational Software Packages
  (Deliverable D4.2)
\item (Month 34) Final Report on the OpenMath Vocabulary
\end{itemize}
\end{wpbox}
\end{workpackage}
\newpage
%%%%%%%%%%%%%% WP 6
\begin{workpackage}{36}{6}{Project Start}{NAG, RISC, BATH}{OpenMath and the Semantic Web}
\begin{wpbox}{Objectives}
???????
\end{wpbox}
\begin{wpbox}{Description of work}
  ????????
\end{wpbox}
\begin{wpbox}{Deliverables\\}
{\bf D6.1} \\
{\bf D6.2} \\
{\bf D6.3} 
\end{wpbox}
\begin{wpbox}{Milestones and expected result}
\begin{itemize}
\item (Month 12) ????
\item (Month 24) ???
\item (Month 36) ???
\end{itemize}
\end{wpbox}
\end{workpackage}
\newpage

%%%%%%%%%%%%%% WP 7
\begin{workpackage}{36}{7}{Project Start}{RISC, BATH}{Mathematical Services in e-Science}
\begin{wpbox}{Objectives}
  ???????
\end{wpbox}
\begin{wpbox}{Description of work}
  ????????
\end{wpbox}
\begin{wpbox}{Deliverables\\}
{\bf D7.1} \\
{\bf D7.2} \\
{\bf D7.3} 
\end{wpbox}
\begin{wpbox}{Milestones and expected result}
\begin{itemize}
\item (Month 12) ????
\item (Month 24) ???
\item (Month 36) ???
\end{itemize}
\end{wpbox}
\end{workpackage}

\newpage
%\setcounter{tocdepth}{2}\tableofcontents

\section{NOT PART OF THE PROPOSAL:  Keep In Mind}
Co-ordination actions are meant to complement other Framework Program
instruments in contributing toward integrating research at European
level through well-planned networking or coordination activities.

This means, we need to list to which projects we are adding bells and
whistles.

GOAL: This document should become the big-picture visionary paper to
rally the OM crowd around for the next part of the OM initiative.  We
need a \emph{new vision} that can be sold to the OM crowd (target 1)
and funding bodies (target 2).


\section{NOT PART OF THE PROPOSAL: Work Plan}
We define a work plan containing a whole range of medium- to long-term
types of networking or co-ordination activities.
\begin{itemize}
\item Performance of studies, analyzes, benchmarking exercises;

test sets for  solving simple problems phrased in OM


\item exchange and dissemination of information

\item exchange and dissemination of good practices

exchanging phrasebooks from one package to another.

Sharing OCDs.
Test each on 3 applications / software systems

\item exchanges of personnel;

offer young researchers for writing phrasebooks;

(small  budget)


\item organization of conference, seminar and meetings;

OpenMath workshop series, and satellites for ISSAC or similar, and
possibly further out, eg economics.

Combine with MathML?

\item setting up of common information systems;

all the data base, knowledge base repository of OCDs

connections with MKM?

Expand and maintain www.openmath.org at NAG


\item setting up of expert groups;

one of these: makers of CAS systems, involved in phrasebooks for
particular sets of OCDs.

$\surd$ editorial board for OCD submission, editing, validation, etc.

\item definition, organization and management of joint or common
  initiatives.

GAP-MAGMA for matrix recognition project

OpenMath for finance,

$\surd$ for chemistry, 

$\surd$ GRID, 

$\surd$ e-learning (LeAM) 

FORM

planar geometry: Cinderella, Botana, ...



\end{itemize}


\end{document}


% LocalWords:  mega Calculemus ednote EdNote newpart BegNP EndNP EdNotes pt tel
% LocalWords:  OpenMathTN WebALT LeActivemath MathBroker GENSS Stilo amc CDs MK
% LocalWords:  ORCCA Eindhoven LeActiveMath Aldor CoCoA TPS PVS Phrasebooks CMU
% LocalWords:  NIST DocBook IST eLearning WebAlt IMACS ACA CML co STMML RoboML
% LocalWords:  Wiris MKM MSDL OM phrasebooks OCDs eg CAS OCD LeAM Botana FET TU
% LocalWords:  SME INRIA Antipolis INLINE Zuse Zentrum Informationstechnik ZIB
% LocalWords:  Explo Centre DFKI Programme DLMF MiKo macs mo workpackage fmpage
% LocalWords:  wpbox
